---
title: "Data exploratory analysis"
author: "Antonia Chroni for St. Jude Children's Research Hospital"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    df_print: paged
    code_folding: hide
    toc_depth: 6
    highlight: tango
    number_sections: TRUE
  pdf_document:
    toc: TRUE
    highlight: tango
    number_sections: TRUE
    latex_engine: lualatex
    keep_tex: FALSE
always_allow_html: TRUE
urlcolor: blue
linkcolor: black
citecolor: blue
geometry: margin=1in
header-includes: 
  - \usepackage{titling}
  - \usepackage{fancyhdr}
params:
  root_dir: './'
  metadata_dir: './'
  PROJECT_NAME: './'
  PI_NAME: './'
  TASK_ID: './'
  PROJECT_LEAD_NAME: './'
  DEPARTMENT: './'
  LEAD_ANALYSTS: './'
  GROUP_LEAD: './'
  CONTACT_EMAIL: './'
  PIPELINE: './'
  START_DATE: './'
  COMPLETION_DATE: './'
---

```{r logo-file, echo=FALSE}
attach(params)
# Insert logo on the top of the html report 
logo_file <- file.path(root_dir, "courses", "Automation-Reproducibility-compbio", "demo", "figures", "img", "DNB-BINF-Core-logo.png")
htmltools::img(src = knitr::image_uri(logo_file), alt = "logo", style = "position:absolute; top:0; left:0; padding:0px; height:120px;")
detach(params)
```

\addtolength{\headheight}{2.0cm} 
\fancypagestyle{plain}{} 
\thispagestyle{fancy}
\fancyhead[L]{\includegraphics[height=120px]{`r logo_file`}}
\renewcommand{\headrulewidth}{0pt}

<style type="text/css">
:root { --DNB_BINF_Core_color: #00427B;}

h1.title {margin-top: 130px;
  margin-bottom: 25px;
  font-size: 36px;}

.nobullet li {list-style-type: none;}

.reporthead {font-size: 20px;}

body { /* Normal */
  font-size: 16px;
  font-style: Arial, Helvetica, sans-serif;}

h1 {
  color: var(--DNB_BINF_Core_color);
  font-size: 28px;
  margin-top: 50px;}

h2 {color: var(--DNB_BINF_Core_color);
  font-size: 20px;}

.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
  background-color: var(--DNB_BINF_Core_color);}
</style>

<a href="https://wiki.stjude.org/display/CAB">

</a>

\pagebreak

<div class="reporthead"><br/>
**PI: `r params$PI_NAME`**  
**Project: `r params$PROJECT_NAME`**  
Task: `r params$TASK_ID`  
Project Lead(s): `r params$PROJECT_LEAD_NAME`  
Department: `r params$DEPARTMENT`  

<br />  

DNB Bioinformatics Core Analysis Team: 
<br />  

>**Lead Analyst(s): `r params$LEAD_ANALYSTS`**  
>Group Lead: `r params$GROUP_LEAD`  
<br />
>**Contact E-mail:** `r params$CONTACT_EMAIL`  
>**DNB Bioinformatics Core Pipeline:** `r params$PIPELINE`  

Date started: `r params$START_DATE`  
Date completed:  `r params$COMPLETION_DATE`  
Report generated: `r format(Sys.time(), '%H:%M:%S %Z %m/%d/%Y')` \

Reviewed by: _____________________   Date: ____________ \
</div>
\pagebreak

# Information about this notebook
This is an exploratory analysis of the data in the project. We are investigating the number of samples per `Sex`. 
This is a demo designed for the purposes of the `Best practices` training in the context of the `Automation and Reproducibility in Computational Biology` course. 


# Set up

```{r load-library, echo=TRUE}
suppressPackageStartupMessages({
  library(tidyverse)
})
```

# Directories and paths to file Inputs/Outputs

```{r set-dir-and-file-names, echo=TRUE}
attach(params)
demo_dir <- file.path(root_dir, "courses", "Automation-Reproducibility-compbio", "demo")
analysis_dir <- file.path(demo_dir, "analyses", "data-exploratory-analysis") 

# We will first read in metadata file as we need to define sample_name 
metadata_file <- file.path(metadata_dir, "project_metadata.tsv") # metadata input file
palette_file <- file.path(demo_dir, "figures", "palettes", "binary_color_palette.tsv")

# File path to `plots` directory
plots_dir <- file.path(analysis_dir, "plots") 
if (!dir.exists(plots_dir)) {
  dir.create(plots_dir)}

figures_plots_dir <- file.path(plots_dir, "figures") 
if (!dir.exists(figures_plots_dir)) {
  dir.create(figures_plots_dir)}

source(paste0(analysis_dir, "/util/function-create-barplot.R"))
source(paste0(demo_dir, "/figures/scripts/theme_plot.R"))
```

# Read metadata file

```{r read-metadata, echo=TRUE}
# Read metadata
df <- read.csv(metadata_file, sep = '\t', header = TRUE) 
```

## Color palette for plotting

```{r palette, echo=TRUE}
# Read color palette
palette_df <- readr::read_tsv(palette_file, guess_max = 100000, show_col_types = FALSE) %>%
  mutate(color_names = case_when(grepl("binary_1", color_names) ~ "M",
                                     grepl("binary_2", color_names) ~ "F"))
       
# Define and order palette
palette <- palette_df$hex_codes
names(palette) <- palette_df$color_names
```

# Number of samples by Sex

```{r, fig.align = "left", results = "asis", message = FALSE, warning = FALSE, echo = FALSE}
tables1 <- df %>% count(Sex)
cat("  \n<div style=\"font-size:80%\">  \n")
caption_value = "Summary of samples per Sex"

print(knitr::kable(tables1, align = "lcccc", caption = caption_value))
cat("  \n</div>  \n")
cat("  \n\\pagebreak  \n")
```

```{r plot1, fig.width = 6, fig.height = 5, fig.fullwidth = TRUE}
# Define parameters for function
ylim <- max(tables1$n)

# Run function
fname <- paste0(figures_plots_dir, "/", "samples-per-sex.pdf")
p <- create_barplot(plot_df = tables1, 
                    ylim = ylim, 
                    x_value = tables1$Sex, 
                    use_palette = palette, 
                    xtitle ="Type of Sex", 
                    title_value = caption_value)
pdf(file = fname, width = 6, height = 5)
print(p)
dev.off()
```

```{r echo=FALSE}
detach(params)
```

\pagebreak

# Session Info

```{r echo=FALSE}
sessionInfo()
```

